---
title: "Getting Started with LibeRties"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with LibeRties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**LibeRties** (Task Infrastructure and Execution Service) runs remote
**LibeRation** jobs with API tokens, per-user limits, and an admin Shiny UI.

## Installation

```r
install.packages(c("plumber", "jsonlite", "digest", "sodium"))
# Install LibeRation + LibeRtAD first, then:
# devtools::install("path/to/LibeRties")
```

## Configuration

Set the sandbox root before starting services (jobs and user data live here):

```{r config, eval = FALSE}
library(LibeRties)

ls_config_set(
  sandbox_root = "C:/liberties/sandbox",
  launcher = "local"   # or "docker"
)
ls_sandbox_root()
```

Environment variable `LIBERTIES_SANDBOX_ROOT` overrides the default. Restart
`ls_run_api()` and `ls_run_admin()` after changing the sandbox.

## Admin token and users

```{r users, eval = FALSE}
ls_admin_token_set("change-me-admin")

u <- ls_user_create("alice", limits = list(
  max_concurrent_jobs = 2,
  max_disk_mb = 5120,
  max_cpu = 4,
  max_memory_mb = 8192
))
u$token   # save — shown once
```

List users and adjust limits from R or the admin Shiny app:

```{r user-mgmt, eval = FALSE}
ls_user_list()
ls_user_set_limits("alice", max_concurrent_jobs = 4)
```

## Start API and admin UI

```{r run, eval = FALSE}
ls_run_api()    # Plumber REST API (default :8080)
ls_run_admin()  # Shiny admin (default :8081)
```

For production, bind the API to `127.0.0.1` and put TLS in front using the
Caddy/nginx examples under `inst/deploy/`.

## LibeRation client

From a workstation with **LibeRation** installed:

```{r client, eval = FALSE}
library(LibeRation)

nm_remote_server_add(
  "HPC", "http://cluster:8080", "alice",
  token = u$token
)
job <- nm_job_submit(model, data, method = "FO", server = "HPC")
nm_job_status(job$id)
fit <- nm_job_result(job$id)
```

The LibeRation Shiny app provides **Add remote server** on the Jobs tab and a
**Run on** selector in the estimation dialog.

## Docker worker

Build the worker image from the repository root:

```bash
docker build -f LibeRties/inst/docker/Dockerfile -t liberties-worker:latest .
```

Then set `launcher = "docker"` in [ls_config_set()].

## Next steps

- API reference: `?ls_run_api`, `?ls_user_create`, `?ls_config`
- Security notes: `SECURITY.md`
- Deployment: `inst/deploy/RUNBOOK.md`
