---
title: "Persistent local and remote job queues"
author: "Sven C. van Dijkman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Persistent local and remote job queues}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Job contract

LibeRties executes LibeRation simulations and estimations through a versioned,
typed job contract. A job contains a serializable `nm_model`, event data,
named entry-point arguments, and an optional label. It must never contain a
live `NMEngine` external pointer.

```{r capabilities}
library(LibeRties)
ls_queue_capabilities()
```

Remote transport uses typed JSON. Model expression IR is recompiled from
semantic fields by the receiving host; the server does not deserialize an RDS
payload supplied over HTTP and does not expose an arbitrary-code endpoint.

## Create a local queue

```{r queue}
queue <- ls_local_queue(
  root = tempfile("liberties-queue-"),
  user = "local",
  max_workers = 1L,
  limits = list(
    max_runtime_seconds = 3600,
    max_memory_mb = 4096,
    max_queued_jobs = 100
  )
)
queue
queue$list()
```

The queue root contains durable job metadata, checksums, payloads, results,
and log streams. Reopening the same root recovers terminal history and checks
whether previously running worker processes are still alive.

## Submit and monitor work

```{r submit, eval=FALSE}
job <- ls_job(
  "estimate",
  model = model,
  data = data,
  arguments = list(
    method = "FOCEI",
    covariance = TRUE,
    print_every = 10,
    n_cores = 2
  ),
  label = "FOCEI base model"
)

id <- queue$submit(job)
queue$status(id)
queue$logs(id, "stdout")
final <- queue$wait(id, timeout = 3600)

if (identical(final$status, "completed")) {
  fit <- queue$result(id)
}
```

States are `queued`, `running`, `completed`, `failed`, and `cancelled`.
`$poll()` reaps workers, enforces limits, recovers untracked jobs, and starts
queued work when capacity is available. `$cancel()` terminates tracked work
and durably records cancellation.

## Integrity and provenance

Payload and result SHA-256 digests are recorded in metadata and verified on
read. Each worker runs in a fresh R process with a scrubbed environment and a
per-job working directory. Runtime, CPU, memory, result-size, queue, and
storage limits are applied at the queue boundary.

This is application-level isolation. A production multi-user host should add
OS-account or container isolation where the threat model requires protection
from native-code compromise.

## Connect to a remote service

```{r remote, eval=FALSE}
remote <- ls_remote(
  "https://liberties.example.org",
  token = Sys.getenv("LIBERTIES_TOKEN")
)

remote$authenticate()
id <- remote$submit(job)
remote$status(id)
remote$logs(id, "stdout")
fit <- remote$result(id)
```

The remote client mirrors the local operations closely. HTTPS is strongly
recommended outside a trusted private network because bearer tokens authorize
all user operations.
