---
title: "Server administration and deployment"
author: "Sven C. van Dijkman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Server administration and deployment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Persistent server root

The server root owns the user registry and per-tenant job namespaces. Choose a
stable directory outside the installed R package. The admin GUI reads
`LIBERTIES_ROOT` or `options(LibeRties.root = ...)` when no explicit root is
supplied.

```{r root}
library(LibeRties)
root <- tempfile("liberties-server-")
server <- ls_server(root, max_workers_per_user = 2L)
server$root
```

## User lifecycle

`ls_user_create()` returns a bearer token once. Only its digest is stored, so
the plaintext token must be transferred and retained securely.

```{r users}
user <- ls_user_create(
  root,
  username = "alice",
  first_name = "Alice",
  last_name = "Example",
  limits = list(
    max_concurrent_jobs = 2,
    max_runtime_seconds = 7200,
    max_memory_mb = 8192,
    max_storage_mb = 20480
  )
)

token <- user$token
ls_user_list(root)[, c("username", "first_name", "last_name", "enabled")]
```

```{r update-users, eval=FALSE}
ls_user_update(root, "alice", enabled = FALSE)
new_token <- ls_user_rotate_token(root, "alice")
ls_user_delete(root, "alice", remove_jobs = FALSE)
```

Rotation invalidates the previous token. Deleting a user is refused while
jobs are active; job history is retained unless `remove_jobs = TRUE` is
explicitly requested.

## Start the HTTP API

```{r api, eval=FALSE}
ls_run_api(
  root = root,
  host = "127.0.0.1",
  port = 8000,
  max_workers_per_user = 2
)
```

Loopback is the secure default. A client on the same machine can connect
directly. Remote clients need a reachable address and firewall rule. A
reverse proxy is not technically required, but it is the recommended way to
provide TLS, request limits, hardened timeouts, access logging, and certificate
renewal.

Only proxies listed in `ls_security_policy(trusted_proxies = ...)` may supply a
forwarded client address. Production preflight also requires a real isolation
probe (or detected Linux container/cgroup evidence); an environment label is
not accepted as proof of a sandbox.

For example, expose `https://liberties.example.org` through a maintained proxy
and forward only to `127.0.0.1:8000`. Do not bind an unencrypted bearer-token
service to a public network.

## Administration GUI

```{r admin, eval=FALSE}
Sys.setenv(LIBERTIES_ROOT = root)
Sys.setenv(LIBERTIES_ADMIN_TOKEN = "a-separate-admin-secret")
ls_run_admin(host = "127.0.0.1")
```

The GUI provides searchable user administration, selectable job tables,
worker logs, state-specific runtime counts, and light/dark themes. User and
job tables are read from the persistent server root; restarting or
reinstalling the package must not erase them.

The admin token protects the administration application and must be distinct
from user bearer tokens. Place the admin GUI behind stronger network access
controls for production use.

## Operational recommendations

- Back up the server root consistently and test restoration.
- Keep LibeRtAD, LibeRation, and LibeRties versions aligned across clients and
  workers.
- Run the service under a dedicated least-privilege OS account.
- Add container or OS-user isolation for mutually untrusted tenants.
- Terminate TLS at a maintained reverse proxy and restrict request sizes.
- Monitor disk, memory, worker failures, queue latency, and repeated
  authentication failures.
- Retain manifests, package versions, logs, and result checksums with modelling
  records.

LibeRties supplies execution durability and application-level tenancy. Formal
validation, regulated-computing controls, disaster-recovery qualification,
electronic signatures, and submission governance remain deployment-specific
responsibilities.
